---
title: "CustomerServiceSTAT"
author: "City of New Orleans, Office of Performance and Accountability"
date: "July 7, 2016"
output: ioslides_presentation
---

```{r setup, include=FALSE}

# Set global settings for markdown document
knitr::opts_chunk$set(echo = FALSE,
                      message = FALSE,
                      warning = FALSE)

# Set working directory
setwd("O:/Projects/Customer Service and Permitting/CustomerServiceSTAT/Script Development")

# Load dependencies
library(devtools)
library(dplyr)
library(ggplot2)
library(grDevices)
library(grid)
library(gridExtra)
library(htmlTable)
library(readxl)
library(RODBC)
library(scales)
library(tidyr)
library(zoo)

theme_opa <- function(base_size = 11, base_family = "") {

    half_line <- base_size/2
    
    theme(

        line = element_line(colour = "black", size = 0.5, linetype = 1, lineend = "butt"),
        rect = element_rect(fill = "white", colour = "black", size = 0.5, linetype = 1),
        text = element_text(family = base_family, face = "plain", colour = "black", size = base_size, lineheight = 0.9, hjust = 0.5, vjust = 0.5, angle = 0, margin = margin(), debug = FALSE),
        
        axis.line = element_line(),
        axis.line.x = element_blank(), 
        axis.line.y = element_blank(),
        axis.text = element_text(size = rel(0.8), colour = "black"),
        axis.text.x = element_text(angle = 45, margin = margin(t = 0.8 * half_line/2), hjust = 1, vjust = 1),
        axis.text.y = element_text(margin = margin(r = 0.8 * half_line/2), hjust = 1),
        axis.ticks = element_blank(),
        axis.ticks.length = unit(half_line/2, "pt"),
        axis.title.x = element_blank(),
        axis.title.y = element_blank(), 
        
        legend.background = element_rect(colour = NA),
        legend.margin = unit(0.2, "cm"),
        legend.key = element_rect(colour = "white"),
        legend.key.size = unit(1.2, "lines"), 
        legend.key.height = NULL,
        legend.key.width = NULL,
        legend.text = element_text(size = rel(0.8)),
        legend.text.align = NULL,
        legend.title = element_text(hjust = 0), 
        legend.title.align = NULL,
        legend.position = "right", 
        legend.direction = NULL,
        legend.justification = "center", 
        legend.box = NULL,
        
        panel.background = element_blank(),
        panel.border = element_blank(),
        panel.grid.major = element_line(colour = "black"),
        panel.grid.major.x = element_blank(),
        panel.grid.major.y = element_line(colour = "black"),
        panel.grid.minor = element_blank(), 
        panel.margin = unit(half_line, "pt"),
        panel.margin.x = NULL, 
        panel.margin.y = NULL,
        panel.ontop = FALSE,
        
        strip.background = element_rect(fill = "black", colour = NA),
        strip.text = element_text(colour = "black", size = rel(0.8)),
        strip.text.x = element_text(margin = margin(t = half_line, b = half_line)),
        strip.text.y = element_text(angle = -90, margin = margin(l = half_line, r = half_line)),
        strip.switch.pad.grid = unit(0.1, "cm"),
        strip.switch.pad.wrap = unit(0.1, "cm"),
        plot.background = element_rect(colour = "white"), 
        plot.title = element_text(hjust = 0, size = rel(1.2), margin = margin(b = half_line * 1.2)),
        plot.margin = margin(half_line, half_line, half_line, half_line),
        complete = TRUE)
    
}
        
theme_set(theme_opa())

stat_colors <- colorRampPalette(c("#005983", "#72d0ff"))

```

## Action items

```{r action_items, results = 'asis'}

# Read data
action_items <- read_excel("action_items.xlsx")

# Coerce date to character
action_items$Assigned <- as.character(action_items$Assigned)

# Insert HTML tag for line breaks
action_items$Responsible <- gsub(";", "<br/>", action_items$Responsible)

# Generate table
htmlTable(action_items,
          rnames = FALSE,
          align = "rlll",
          align.header = "rlll",
          css.cell = "vertical-align: top; font-size: 15px; padding: 6px")

```

## 311

---

311 call volume appears to be on pace with prior years

```{r call_volume_311}

# Load and clean 311 performance metrics

internal_311 <- read_excel(path = "311.xlsx", sheet = 1)

names(internal_311) <- tolower(names(internal_311))
internal_311$measure <- tolower(internal_311$measure)

# Collapse date columns into one variable
internal_311 <- gather(internal_311, date, value, 3:ncol(internal_311))

internal_311$value <- as.numeric(internal_311$value)
internal_311$date <- as.yearmon(internal_311$date)

internal_311 <- internal_311 %>% filter(date >= '2014-05-01')

# Generate plot

measure_type <- "calls"

plot_data <- internal_311 %>% filter(measure == measure_type)

ggplot(plot_data, aes(x = date, y = value)) +
    geom_bar(stat = "identity", fill = "#005983") +
    ggtitle("311: Call volume") +
    xlab("Month") +
    ylab("Calls") +
    scale_x_yearmon(breaks = plot_data$date)

```

Source: Office of Information Technology and Innovation

---

Call abandonment rates have remained below two percent since October

```{r abandonment_rate_311}

# Generate plot

measure_type <- "abandonment rate"

plot_data <- internal_311 %>% filter(measure == measure_type)

target <- 8

ggplot(plot_data, aes(x = date, y = value * 100)) +
    geom_line(size = 1.5, color = "#005983") +
    geom_line(aes(y = target), linetype = 2) +
    ggtitle("311: Call abandonment rate") +
    xlab("Month") +
    ylab("Percent") +
    scale_x_yearmon(breaks = plot_data$date) + expand_limits(y = 0)

```

Source: Office of Information Technology and Innovation

---

Hold times remain below 10 seconds

```{r hold_time_311}

# Generate plot

measure_type <- "avg. hold time (sec)"

plot_data <- internal_311 %>% filter(measure == measure_type)

ggplot(plot_data, aes(x = date, y = value)) +
    geom_line() +
    xlab("Month") +
    ylab("Seconds") +
    geom_line(size = 1.5, color = "#005983") +
    scale_x_yearmon(breaks = plot_data$date) +
    ggtitle("311: Average hold time") +
    expand_limits(y = 0) +
    scale_y_continuous(breaks = pretty_breaks(5))

```

Source: Office of Information Technology and Innovation

---

First-call resolution rates dipped below 60 percent in May

```{r first_call_resolution_311}

# Generate plot

measure_type <- "first call resolution"

plot_data <- internal_311 %>% filter(measure == measure_type)

target = 65

ggplot(plot_data, aes(x = date, y = value * 100)) +
    geom_line(size = 1.5, color = "#005983") +
    xlab("Month") +
    ylab("Percent") +
    scale_x_yearmon(breaks = plot_data$date) +
    ggtitle("311: First-call resolution rate") +
    expand_limits(y = 0) +
    scale_y_continuous(breaks = pretty_breaks(6))

```

Source: Office of Information Technology and Innovation

---

One operator had a total score below 75 in May

```{r operators_311}

# Load and clean operator performance metrics

operators <- read_excel(path = "311.xlsx", sheet = 2, skip = 1)

names(operators) <- tolower(gsub(" ", "_", names(operators)))
names(operators) <- gsub("_score.*", "", names(operators))

operators <- operators %>% select(-customer_survey) %>% mutate(total = total * 100)

operators <- gather(operators, metric, value, 2:ncol(operators))

operators <- operators %>%
    mutate(operator = gsub("Operator ", "", agent)) %>%
    select(operator, metric, value)

operators$metric <- relevel(factor(gsub("_", " ", operators$metric)), "total")

# Plot operator performance

ggplot(operators, aes(x = factor(operator, levels = 1:12), y = value, fill = metric)) +
    geom_bar(stat = "identity", position = "dodge") +
    ggtitle("Operator scores (out of 100)") +
    labs(fill = NULL) +
    xlab("Operator") +
    ylab("Score") +
    theme(plot.title = element_text(hjust = 0)) +
    scale_fill_manual(values = stat_colors(6)) +
    theme(axis.text.x = element_text(angle = 0))

```

Source: Office of Information Technology and Innovation

## Permitting

---

Online permit applications have been steadily increasing

```{r sp_permit_applications}

channel <- odbcDriverConnect(connection = "Driver={SQL Server};Server=cno-sqlreport01;Database=Lama_Rpt")
    
query <- paste(readLines(con = "P:/Safety and Permits/sp_permit_applications.sql"), collapse = "\t")
    
# Query database
permit_applications <- sqlQuery(channel, query, as.is = TRUE)
    
# Close database connection
close(channel)

names(permit_applications) <- tolower(names(permit_applications))

permit_applications <- permit_applications %>%
    mutate(datesubmitted = as.Date(datesubmitted),
           m = as.yearmon(datesubmitted),
           online = factor(online))

levels(permit_applications$online) <- c("In person", "Online")

plot_data <- permit_applications %>%
    group_by(online, m) %>%
    summarize(total = n())

# Generate plot
ggplot(data = plot_data, aes(x = m, y = total)) +
    geom_bar(aes(fill = factor(online)), position = "dodge", stat = "identity") +
    scale_x_yearmon(breaks = plot_data$m) +
    labs(fill = NULL) +
    ggtitle("Safety and Permits: Applications submitted by month") + scale_fill_manual(values = stat_colors(2))

```

Source: LAMA; OPA database query

---

The percentage of online applications continues to increase, due in part to declining in-person applications

```{r sp_percent_online}

# Filter for application types that can be submitted online
plot_data <- permit_applications %>%
    filter(iswebapp == 1) %>%
    group_by(m) %>%
    summarize(perc_online = sum(online == "Online")/n() * 100)

# Generate plot
ggplot(data = plot_data, aes(x = m, y = perc_online)) +
    geom_line(size = 1.5, color = "#005983") +
    scale_x_yearmon(breaks = plot_data$m) +
    labs(color = NULL) +
    ylab("Percent") +
    ggtitle("Safety and Permits: Percent of permit applications submitted online") +
    expand_limits(y = 0) +
    scale_y_continuous(breaks = pretty_breaks(7))

```

Source: LAMA; OPA database query

---

Commercial permit issuances have been steady, while the number of residential permits has shown a slight decrease

```{r sp_res_comm_issued}

channel <- odbcDriverConnect(connection = "Driver={SQL Server};Server=cno-sqlreport01;Database=Lama_Rpt")
    
query <- paste(readLines(con = "P:/Safety and Permits/sp_permits_issued.sql"), collapse = "\t")
    
# Query database
permits_issued <- sqlQuery(channel, query, as.is = TRUE)
    
# Close database connection
close(channel)

names(permits_issued) <- tolower(names(permits_issued))

permits_issued <- permits_issued %>%
    mutate(datesubmitted = as.Date(datesubmitted),
           issuedate = as.Date(issuedate),
           m = as.yearmon(issuedate))

# Move "Residential" permits to first position on legend
permits_issued$use <- relevel(factor(permits_issued$use), "Residential")

plot_data <- permits_issued %>%
    filter(use %in% c('Residential', 'Commercial')) %>%
    group_by(use, m) %>%
    summarize(total = n())

ggplot(data = plot_data, aes(x = m, y = total)) +
    geom_bar(aes(fill = use), position = "dodge", stat = "identity") +
            scale_x_yearmon(breaks = unique(permits_issued$m)) +
            labs(fill = "Land use") +
            xlab("Month") +
            ylab("Permits") +
            ggtitle("Safety and Permits: Permits issued by month") + scale_fill_manual(values = stat_colors(2))

```

Source: LAMA; OPA database query

---

Time to issue commercial and residential permits has increased slightly over the past two years

```{r sp_res_comm_days_to_issue}

permits_issued$use <- relevel(factor(permits_issued$use), "Commercial")

daystoissue <- permits_issued %>%
    filter(use %in% c('Residential', 'Commercial')) %>%
    filter(daystoissue >= 0) %>%
    group_by(use, m) %>%
    summarize(mean = mean(daystoissue))

plot_data <- daystoissue

ggplot(data = plot_data, aes(x = m, y = mean, color = factor(use))) +
    geom_line(size = 1.5) +
    scale_x_yearmon(breaks = plot_data$m) +
    labs(color = NULL) +
    xlab("Month") +
    ylab("Days") +
    ggtitle("Safety and Permits: Mean days to issue permit") +
    scale_color_manual(values = stat_colors(2)) +
    expand_limits(y = 0) +
    scale_y_continuous(breaks = pretty_breaks(6))

```

Source: LAMA; OPA database query

---

Issuance of event and unknown/other permit types has risen over the past two years

```{r sp_event_issued}

# Move "Residential" permits to first position on legend
permits_issued$use <- relevel(factor(permits_issued$use), "Event")

plot_data <- permits_issued %>%
    filter(!(use %in% c('Residential', 'Commercial'))) %>%
    group_by(use, m) %>%
    summarize(total = n())

ggplot(data = plot_data, aes(x = m, y = total)) +
    geom_bar(aes(fill = use), position = "dodge", stat = "identity") +
            scale_x_yearmon(breaks = unique(permits_issued$m)) +
            labs(fill = "Land use") +
            xlab("Month") +
            ylab("Permits") +
            ggtitle("Safety and Permits: Permits issued by month") + scale_fill_manual(values = stat_colors(2))

```

Source: LAMA; OPA database query

---

Time to issue events permits has varied considerably

```{r sp_event_days_to_issue}

permits_issued$use <- relevel(factor(permits_issued$use), "Commercial")

daystoissue <- permits_issued %>%
    filter(!(use %in% c('Residential', 'Commercial'))) %>%
    filter(daystoissue >= 0) %>%
    group_by(use, m) %>%
    summarize(mean = mean(daystoissue))

plot_data <- daystoissue

ggplot(data = plot_data, aes(x = m, y = mean, color = factor(use))) +
    theme(axis.text.x = element_text(angle = 45, hjust = 1),
    plot.title = element_text(hjust = 0)) +
    geom_line(size = 1.5) +
    scale_x_yearmon(breaks = plot_data$m) +
    labs(color = NULL) +
    xlab("Month") +
    ylab("Days") +
    ggtitle("Safety and Permits: Mean days to issue permit") +
    scale_color_manual(values = stat_colors(2)) +
    expand_limits(y = 0) +
    scale_y_continuous(breaks = pretty_breaks(7))

```

Source: LAMA; OPA database query

---

Percentage of building permits issued within one day appears to have fallen, but is still above 40 percent

```{r sp_building_permits}

building_permits <- permits_issued %>%
    filter(building == 1) %>%
    group_by(m) %>%
    summarize(perc_in_target = sum(daystoissue <= 1)/n() * 100)

plot_data <- building_permits

ggplot(data = plot_data, aes(x = m, y = perc_in_target)) +
    geom_line(size = 1.5, color = "#005983") +
    scale_x_yearmon(breaks = plot_data$m) +
    labs(color = NULL) +
    ylab("Percent") +
    ggtitle("Safety and Permits: Percentage of building permits issued within one day") +
    expand_limits(y = 0)

```

Source: LAMA; OPA database query

---

Time to complete zoning inspections for business licenses has been within target since November

```{r sp_business_license_inspections}

channel <- odbcDriverConnect(connection = "Driver={SQL Server};Server=cno-sqlreport01;Database=Lama_Rpt")
    
query <- paste(readLines(con = "P:/Safety and Permits/sp_business_license_inspections.sql"), collapse = "\t")
    
# Query database
df <- sqlQuery(channel, query, as.is = TRUE)
    
# Close database connection
close(channel)

license_inspections <- df

names(license_inspections) <- tolower(names(license_inspections))

target <- 7

license_inspections <- license_inspections %>%
    mutate(datefiled = as.Date(datefiled),
           dateinspected = as.Date(dateinspected),
           m = as.yearmon(dateinspected),
           in_target = factor(ifelse(daystoinspect <= target, "Seven days or fewer", "More than seven days")))

license_inspections$in_target <- relevel(license_inspections$in_target, "Seven days or fewer")

plot_data <- license_inspections %>%
    group_by(m) %>%
    summarize(mean = mean(daystoinspect))

ggplot(data = plot_data, aes(x = m, y = mean)) +
    geom_line(size = 1.5, color = "#005983") +
    scale_x_yearmon(breaks = plot_data$m) +
    labs(color = NULL) +
    xlab("Month") +
    ylab("Days") +
    ggtitle("Safety and Permits: Mean days to complete business license zoning inspections") +
    expand_limits(y = 0) +
    scale_y_continuous(breaks = pretty_breaks(5))

```

Source: LAMA; OPA database query

---

The vast majority of business license zoning inspections are completed within seven days

```{r sp_business_license_inspections_by_time}

plot_data <- license_inspections %>%
    group_by(in_target, m) %>%
    summarize(total = n())

ggplot(data = plot_data, aes(x = m, y = total, fill = in_target)) +
    geom_bar(position = "dodge", stat = "identity") +
    scale_x_yearmon(breaks = plot_data$m) +
    labs(fill = NULL) +
    xlab("Month") +
    ylab("Inspections") +
    ggtitle("Safety and Permits: Business license inspections by time elapsed from application") +
    scale_fill_manual(values = stat_colors(2))

```

Source: LAMA; OPA database query

---

Outliers have pushed up the average time to issue HDLC permits, but median time to issue has been less than one day

```{r hdlc_permits_issued}

channel <- odbcDriverConnect(connection = "Driver={SQL Server};Server=cno-sqlreport01;Database=Lama_Rpt")
    
query <- paste(readLines(con = "P:/Historic District Landmarks Commission/hdlc_permits_issued.sql"), collapse = "\t")
    
# Query database
df <- sqlQuery(channel, query, as.is = TRUE)
    
# Close database connection
close(channel)

permits_issued <- df

names(permits_issued) <- tolower(names(permits_issued))

permits_issued <- permits_issued %>%
    mutate(datecreated = as.Date(datecreated),
           dateissued = as.Date(dateissued),
           m = as.yearmon(dateissued))

plot_data <- permits_issued %>%
    group_by(m) %>%
    summarize(mean = mean(daystoissue), median = median(daystoissue))

ggplot(data = plot_data, aes(x = m)) +
    geom_line(aes(y = mean, color = "Mean"), size = 1.5) +
    geom_line(aes(y = median, color = "Median"), size = 1.5) +
    scale_x_yearmon(breaks = plot_data$m) +
    labs(color = NULL) +
    theme(axis.text.x = element_text(angle = 45, hjust = 1), plot.title = element_text(hjust = 0)) +
    xlab("Month") +
    ylab("Days") +
    ggtitle("Historic District Landmarks Commission: Days to issue permit") +
    scale_color_manual(values = stat_colors(2))

```

Source: LAMA; OPA database query

---

The majority of staff-approvable HDLC permits are issued within three days

```{r hdlc_permits_by_time_to_issue}

target <- 3

permits_issued <- permits_issued %>%
    mutate(m = as.yearmon(dateissued),
           in_target = factor(ifelse(daystoissue <= target, "Three days or fewer", "More than three days")))

permits_issued$in_target <- relevel(permits_issued$in_target, "Three days or fewer")

plot_data <- permits_issued %>%
    group_by(in_target, m) %>%
    summarize(total = n())

ggplot(data = plot_data, aes(x = m, y = total, fill = in_target)) +
    theme(axis.text.x = element_text(angle = 45, hjust = 1), plot.title = element_text(hjust = 0)) +
    geom_bar(position = "dodge", stat = "identity") +
    scale_x_yearmon(breaks = plot_data$m) +
    labs(fill = NULL) +
    xlab("Month") +
    ylab("Permits") +
    ggtitle("HDLC: Permits by time to issue") +
    scale_fill_manual(values = stat_colors(2))

```

Source: LAMA; OPA database query

---

VCC staff have been working to update the LAMA event flow for permit issuance

```{r vcc_permits_issued}

channel <- odbcDriverConnect(connection = "Driver={SQL Server};Server=cno-sqlreport01;Database=Lama_Rpt")
    
query <- paste(readLines(con = "P:/Vieux Carre Commission/vcc_permits_issued.sql"), collapse = "\t")
    
# Query database
df <- sqlQuery(channel, query, as.is = TRUE)
    
# Close database connection
close(channel)

permits_issued <- df

names(permits_issued) <- tolower(names(permits_issued))

permits_issued <- permits_issued %>%
    mutate(datesubmitted = as.Date(datesubmitted),
           dateissued = as.Date(dateissued),
           m = as.yearmon(dateissued))

plot_data <- permits_issued %>%
    group_by(m) %>%
    summarize(mean = mean(daystoissue))

ggplot(data = plot_data, aes(x = m, y = mean)) +
    geom_line(size = 1.5, color = "#005983") +
    scale_x_yearmon(breaks = plot_data$m) +
    labs(color = NULL) +
    ggtitle("Vieux Carre Commission: Mean days to issue staff-approvable permits") +
    expand_limits(y = 0)

```

Source: LAMA; OPA database query

---

The vast majority of staff-approvable VCC permits are also issued within three days

```{r vcc_permits_by_time_to_issue}

target <- 3

permits_issued <- permits_issued %>%
    mutate(m = as.yearmon(dateissued),
           in_target = factor(ifelse(daystoissue <= target, "Three days or fewer", "More than three days")))

permits_issued$in_target <- relevel(permits_issued$in_target, "Three days or fewer")

plot_data <- permits_issued %>%
    group_by(in_target, m) %>%
    summarize(total = n())

ggplot(data = plot_data, aes(x = m, y = total, fill = in_target)) +
    theme(axis.text.x = element_text(angle = 45, hjust = 1), plot.title = element_text(hjust = 0)) +
    geom_bar(position = "dodge", stat = "identity") +
    scale_x_yearmon(breaks = plot_data$m) +
    labs(fill = NULL) +
    xlab("Month") +
    ylab("Permits") +
    ggtitle("VCC: Permits by time to issue") +
    scale_fill_manual(values = stat_colors(2))

```

Source: LAMA; OPA database query

## One Stop Shop

---

Median wait time for any license or permit queue has been above target since January but is trending back down

```{r oss}

channel <- odbcDriverConnect(connection = "Driver={SQL Server};Server=CNO-SQLCST01;Database=lobbyCentral")
    
query <- paste(readLines(con = "P:/Safety and Permits/oss_wait_service_times.sql"), collapse = "\t")
    
# Query database
df <- sqlQuery(channel, query, as.is = TRUE)
    
# Close database connection
close(channel)

oss <- df

names(oss) <- tolower(names(oss))
oss$queue <- tolower(oss$queue)

oss <- oss %>%
    filter(waittime >= 0 & waittime <= 480,
           servicetime >= 0 & servicetime <= 480) %>%
    mutate(arrived = as.POSIXct(arrived),
           assisted = as.POSIXct(assisted),
           departed = as.POSIXct(departed),
           m = as.yearmon(arrived))

# Consolidate queues into larger categories if needed
taxi_queues <- c("brake tag",
                 "bus permit",
                 "cpnc",
                 "driver/operator",
                 "taxi deputy director")

# Calculate performance measures

any_lic_perm_wait <- oss %>%
    filter(grepl("license|permit", queue)) %>%
    group_by(m) %>%
    summarize(med_wait_time = median(waittime), med_service_time = median(servicetime), customers = n())

build_perm_wait <- oss %>%
    filter(queue == "building permit") %>%
    group_by(m) %>%
    summarize(med_wait_time = median(waittime), med_service_time = median(servicetime), customers = n())

occ_lic_wait <- oss %>%
    filter(queue == "occup. license") %>%
    group_by(m) %>%
    summarize(med_wait_time = median(waittime), med_service_time = median(servicetime), customers = n())

payment_wait <- oss %>%
    filter(queue == "payment") %>%
    group_by(m) %>%
    summarize(med_wait_time = median(waittime), med_service_time = median(servicetime), customers = n())

taxi_wait <- oss %>%
    filter(queue %in% taxi_queues) %>%
    group_by(m) %>%
    summarize(med_wait_time = median(waittime), med_service_time = median(servicetime), customers = n())

```

```{r oss_any_lic_perm}

lobby_central_plot <- function(x, title) {

plot_data <- x

p1 <- ggplot(data = plot_data, aes(x = m, y = med_wait_time)) +
    geom_line(size = 1.5, color = "#005983") +
    scale_x_yearmon(breaks = plot_data$m) +
    labs(color = NULL) +
    ggtitle("Median wait time (minutes)") +
    scale_y_continuous(breaks = pretty_breaks(4))

p2 <- ggplot(data = plot_data, aes(x = m, y = med_service_time)) +
    geom_line(size = 1.5, color = "#005983") +
    scale_x_yearmon(breaks = plot_data$m) +
    labs(color = NULL) +
    ggtitle("Median service time (minutes)") +
    scale_y_continuous(breaks = pretty_breaks(4))

p3 <- ggplot(data = plot_data, aes(x = m, y = customers)) +
    geom_bar(stat = "identity", fill = "#005983") +
    scale_x_yearmon(breaks = unique(plot_data$m)) +
    ggtitle("Customers")

grid.arrange(p1, p2, p3, ncol = 1, top = title)

}

lobby_central_plot(any_lic_perm_wait, "One Stop Shop: Any license or permit queue")

```

Source: LobbyCentral; OPA database query

---

Median wait time for the building permit queue has been above target since March but is trending back down

```{r oss_building_permit_queue}

lobby_central_plot(build_perm_wait, "One Stop Shop: Building permit queue")

```

Source: LobbyCentral; OPA database query

---

Median wait time for the occupational license queue has been above target since March but is also trending back down

```{r oss_occupational_license_queue}

lobby_central_plot(occ_lic_wait, "One Stop Shop: Occupational license queue")

```

Source: LobbyCentral; OPA database query

---

Payment wait times have recently been within target

```{r oss_payment_queue}

lobby_central_plot(payment_wait, "One Stop Shop: Payment queue")

```

Source: LobbyCentral; OPA database query

## Bureau of Revenue

```{r revenue}

channel <- odbcDriverConnect(connection = "Driver={SQL Server};Server=CNO-SQLCST01;Database=lobbyCentral")
    
query <- paste(readLines(con = "P:/Revenue/revenue_wait_service_times.sql"), collapse = "\t")
    
# Query database
df <- sqlQuery(channel, query, as.is = TRUE)
    
# Close database connection
close(channel)

revenue <- df

names(revenue) <- tolower(names(revenue))
revenue$type <- tolower(revenue$type)

revenue <- revenue %>%
    filter(waittime >= 0 & waittime <= 480,
           servicetime >= 0 & servicetime <= 480) %>%
    mutate(arrived = as.POSIXct(arrived),
           assisted = as.POSIXct(assisted),
           departed = as.POSIXct(departed),
           m = as.yearmon(arrived))

# Calculate performance measures

abo <- revenue %>%
    filter(type == "abo") %>%
    group_by(m) %>%
    summarize(med_wait_time = median(waittime), med_service_time = median(servicetime), customers = n())

accounts <- revenue %>%
    filter(type == "accounts") %>%
    group_by(m) %>%
    summarize(med_wait_time = median(waittime), med_service_time = median(servicetime), customers = n())

administration <- revenue %>%
    filter(type == "administration") %>%
    group_by(m) %>%
    summarize(med_wait_time = median(waittime), med_service_time = median(servicetime), customers = n())

business <- revenue %>%
    filter(type == "business") %>%
    group_by(m) %>%
    summarize(med_wait_time = median(waittime), med_service_time = median(servicetime), customers = n())

enforcement_events <- revenue %>%
    filter(type == "enforcement/special events") %>%
    group_by(m) %>%
    summarize(med_wait_time = median(waittime), med_service_time = median(servicetime), customers = n())

```

---

Wait times for ABO-related queues have been rising, possibly in line with seasonal volume

```{r revenue_abo_queues}

lobby_central_plot(abo, "Revenue: ABO queues")

```

Source: LobbyCentral; OPA database query

---

Median wait time for account-related queues has varied around five minutes

```{r revenue_account_queues}

lobby_central_plot(accounts, "Revenue: Account queues")

```

Source: LobbyCentral; OPA database query

---

Wait times for the business queue have been held to five minutes or less since June 2015

```{r revenue_business_license_queues}

lobby_central_plot(business, "Revenue: Business license queue")

```

Source: LobbyCentral; OPA database query

---

Wait times for the enforcement and special events queue fell in May, possibly due to seasonal fluctuations

```{r revenue_delinquent_events_queue}

lobby_central_plot(enforcement_events, "Revenue: Delinquent taxes and special events queue")

```

Source: LobbyCentral; OPA database query

## Violations

---

Median time to inspect building violations has varied around one day

```{r sp_building_violations}

channel <- odbcDriverConnect(connection = "Driver={SQL Server};Server=cno-sqlreport01;Database=Lama_Rpt")
    
query <- paste(readLines(con = "P:/Safety and Permits/sp_compliance.sql"), collapse = "\t")
    
# Query database
df <- sqlQuery(channel, query, as.is = TRUE)
    
# Close database connection
close(channel)

violation_inspections <- df

names(violation_inspections) <- tolower(names(violation_inspections))

violation_inspections <- violation_inspections %>%
    filter(category != "Other") %>%
    filter(daystoinspect >= 0) %>%
    mutate(filingdate = as.Date(filingdate),
           inspectiondate = as.Date(inspectiondate),
           m = as.yearmon(inspectiondate)) %>%
    group_by(category, m) %>%
    summarize(mean = mean(daystoinspect), median = median(daystoinspect))

plot_data <- violation_inspections %>% filter(category == "Building")
target <- 5

ggplot(data = plot_data, aes(x = m)) +
    geom_line(aes(y = mean, color = "Mean"), size = 1.5) +
    geom_line(aes(y = median, color = "Median"), size = 1.5) +
    scale_x_yearmon(breaks = plot_data$m) +
    labs(color = NULL) +
    theme(axis.text.x = element_text(angle = 45, hjust = 1), plot.title = element_text(hjust = 0)) +
    xlab("Month") +
    ylab("Days") +
    ggtitle("Safety and Permits: Time from creation of building violation case to inspection") +
    scale_color_manual(values = stat_colors(2))

```

Source: LAMA; OPA database query

---

Median time to inspect zoning violations has also varied around one day

```{r sp_zoning_violations}

plot_data <- violation_inspections %>% filter(category == "Zoning")
target <- 7

ggplot(data = plot_data, aes(x = m)) +
    geom_line(aes(y = mean, color = "Mean"), size = 1.5) +
    geom_line(aes(y = median, color = "Median"), size = 1.5) +
    scale_x_yearmon(breaks = plot_data$m) +
    labs(color = NULL) +
    theme(axis.text.x = element_text(angle = 45, hjust = 1), plot.title = element_text(hjust = 0)) +
    xlab("Month") +
    ylab("Days") +
    ggtitle("Safety and Permits: Time from creation of zoning violation case to inspection") +
    scale_color_manual(values = stat_colors(2))

```

Source: LAMA; OPA database query

## Taxi and For-Hire Bureau

---

Vehicle inspections appear to be slightly ahead of prior-year trends for May

```{r taxi_vehicle_inspections}

channel <- odbcDriverConnect(connection = "Driver={SQL Server};Server=cno-sqlreport01;Database=Lama_Rpt")
    
query <- paste(readLines(con = "P:/Taxicab and For-Hire Bureau/vehicle_inspections.sql"), collapse = "\t")
    
# Query database
df <- sqlQuery(channel, query, as.is = TRUE)
    
# Close database connection
close(channel)

vehicle_inspections <- df

names(vehicle_inspections) <- tolower(names(vehicle_inspections))

vehicle_inspections <- vehicle_inspections %>%
    mutate(dateinspected = as.Date(dateinspected),
           m = as.yearmon(dateinspected))
           
plot_data <- vehicle_inspections %>%
    group_by(m) %>%
    summarize(total = n())

# Generate plot
ggplot(data = plot_data, aes(x = m, y = total)) +
    geom_bar(stat = "identity", fill = "#005983") +
    scale_x_yearmon(breaks = plot_data$m) +
    labs(fill = NULL) +
    ggtitle("Taxi and For-Hire Bureau: Vehicle inspections conducted")

```

Note: Licenses generally come up for renewal in June and December.
Source: LAMA; OPA database query

---

Taxi and For-Hire Bureau wait times have been within target since April 2015

```{r oss_taxi_queues}

lobby_central_plot(taxi_wait, "One Stop Shop: Taxi and For-Hire queues")

```

Source: LobbyCentral; OPA database query

## Questions
