---
title: "CustomerServiceSTAT"
author: "City of New Orleans, Office of Performance and Accountability"
date: "July 7, 2016"
output: ioslides_presentation
---

```{r setup, include=FALSE}

knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE, fig.width = 7.5, fig.height = 5.5)

```

```{r}
setwd("O:/Projects/Customer Service and Permitting/CustomerServiceSTAT/Script Development")

library(dplyr)
library(ggplot2)
library(grid)
library(gridExtra)
library(readxl)
library(tidyr)
library(zoo)

# Load and clean 311 performance metrics

internal_311 <- read_excel(path = "311.xlsx", sheet = 1)

names(internal_311) <- tolower(names(internal_311))
internal_311$measure <- tolower(internal_311$measure)

# Collapse date columns into one variable
internal_311 <- gather(internal_311, date, value, 3:ncol(internal_311))

internal_311$value <- as.numeric(internal_311$value)
internal_311$date <- as.yearmon(internal_311$date)

```

## 311

---



```{r}

# Generate plots

measure_type <- "calls"

plot_data <- internal_311 %>% filter(measure == measure_type)

ggplot(plot_data, aes(x = date, y = value)) +
    geom_bar(stat = "identity") +
    scale_x_yearmon() +
    ggtitle("311: Call volume") +
    xlab("Month") +
    ylab("Calls") +
    theme(axis.text.x = element_text(angle = 45, hjust = 1),
                  plot.title = element_text(hjust = 0)) +
            scale_x_yearmon(breaks = plot_data$date)

```

Source: Office of Information Technology and Innovation

---

```{r}

# Generate plots

measure_type <- "abandonment rate"

plot_data <- internal_311 %>% filter(measure == measure_type)

target <- 8

ggplot(plot_data, aes(x = date, y = value * 100)) +
    geom_line() +
    geom_line(aes(y = target), linetype = 2) +
    ggtitle("311: Call abandonment rate") +
    xlab("Month") +
    ylab("Percent") +
    theme(axis.text.x = element_text(angle = 45, hjust = 1),
                  plot.title = element_text(hjust = 0)) +
            geom_line(size = 1.5) +
            scale_x_yearmon(breaks = plot_data$date)

```

Source: Office of Information Technology and Innovation

---

```{r}

measure_type <- "avg. hold time (sec)"

plot_data <- internal_311 %>% filter(measure == measure_type)

ggplot(plot_data, aes(x = date, y = value)) +
    geom_line() +
    xlab("Month") +
    ylab("Seconds") +
    theme(axis.text.x = element_text(angle = 45, hjust = 1),
                  plot.title = element_text(hjust = 0)) +
            geom_line(size = 1.5) +
            scale_x_yearmon(breaks = plot_data$date) +
    ggtitle("311: Average hold time")

```

Source: Office of Information Technology and Innovation

---

```{r}

measure_type <- "first call resolution"

plot_data <- internal_311 %>% filter(measure == measure_type)

target = 65

ggplot(plot_data, aes(x = date, y = value * 100)) +
    geom_line() +
    geom_line(aes(y = target), linetype = 2) +
    xlab("Month") +
    ylab("Percent") +
    theme(axis.text.x = element_text(angle = 45, hjust = 1),
                  plot.title = element_text(hjust = 0)) +
            geom_line(size = 1.5) +
            scale_x_yearmon(breaks = plot_data$date) +
    ggtitle("311: First call resolution rate")

```

Source: Office of Information Technology and Innovation

---

```{r}

# Load and clean operator performance metrics

operators <- read_excel(path = "311.xlsx", sheet = 2, skip = 1)

names(operators) <- tolower(gsub(" ", "_", names(operators)))
names(operators) <- gsub("_score.*", "", names(operators))

operators <- operators %>% select(-customer_survey) %>% mutate(total = total * 100)

operators <- gather(operators, metric, value, 2:ncol(operators))

operators <- operators %>%
    mutate(operator = gsub("Operator ", "", agent)) %>%
    select(operator, metric, value)

operators$metric <- relevel(factor(gsub("_", " ", operators$metric)), "total")

# Plot operator performance

ggplot(operators, aes(x = factor(operator, levels = 1:12), y = value, fill = metric)) +
    geom_bar(stat = "identity", position = "dodge") +
    ggtitle("Operator scores (out of 100)") +
    labs(fill = NULL) +
    xlab("Operator") +
    ylab("Score") +
    theme(plot.title = element_text(hjust = 0))

```

Source: Office of Information Technology and Innovation

## Permitting

---

```{r permit_applications}

permit_applications <- read_excel("permit_applications.xlsx")

names(permit_applications) <- tolower(names(permit_applications))

permit_applications <- permit_applications %>%
    mutate(m = as.yearmon(datesubmitted)) %>%
    mutate(online = factor(online))

levels(permit_applications$online) <- c("In person", "Online")

ggplot(data = permit_applications, aes(x = m)) +
            theme(axis.text.x = element_text(angle = 45, hjust = 1),
                  plot.title = element_text(hjust = 0)) +
            geom_bar(aes(fill = factor(online)), position = "dodge", stat = "bin") +
            scale_x_yearmon(breaks = unique(permit_applications$m)) +
            labs(fill = NULL) +
            xlab("Month") +
            ylab("Applications") +
            ggtitle("Safety and Permits: Applications submitted by month")


```

Source: LAMA; OPA database query

---

```{r percent_online}

plot_data <- permit_applications %>%
    filter(iswebapp == 1) %>%
    group_by(m) %>%
    summarize(perc_online = sum(online == "Online")/n() * 100)

ggplot(data = plot_data, aes(x = m, y = perc_online)) +
            theme(axis.text.x = element_text(angle = 45, hjust = 1),
                  plot.title = element_text(hjust = 0)) +
            geom_line(size = 1.5) +
            scale_x_yearmon(breaks = plot_data$m) +
            labs(color = NULL) +
            xlab("Month") +
            ylab("Percent") +
            ggtitle("Safety and Permits: Percent of permit applications submitted online")


```

Source: LAMA; OPA database query

---

```{r sp_permits_issued}

permits_issued <- read_excel("permits_issued.xlsx")

names(permits_issued) <- tolower(names(permits_issued))

permits_issued <- permits_issued %>%
    mutate(m = as.yearmon(issuedate))

permits_issued$use <- relevel(factor(permits_issued$use), "Residential")

ggplot(data = permits_issued, aes(x = m)) +
            theme(axis.text.x = element_text(angle = 45, hjust = 1),
                  plot.title = element_text(hjust = 0)) +
            geom_bar(aes(fill = use), position = "dodge", stat = "bin") +
            scale_x_yearmon(breaks = unique(permits_issued$m)) +
            labs(fill = NULL) +
            xlab("Month") +
            ylab("Permits") +
            ggtitle("Safety and Permits: Permits issued by month")

```

Source: LAMA; OPA database query

---

```{r sp_days_to_issue}

daystoissue <- permits_issued %>%
    group_by(use, m) %>%
    summarize(mean = mean(daystoissue))

target <- 15

plot_data <- daystoissue

ggplot(data = plot_data, aes(x = m, y = mean, color = factor(use))) +
            theme(axis.text.x = element_text(angle = 45, hjust = 1),
                  plot.title = element_text(hjust = 0)) +
            geom_line(size = 1.5) +
            geom_line(aes(y = target), linetype = 2) +
            scale_x_yearmon(breaks = plot_data$m) +
            labs(color = NULL) +
            xlab("Month") +
            ylab("Days") +
            ggtitle("Safety and Permits: Mean days to issue permit")

```

Source: LAMA; OPA database query

---

```{r sp_building_permits}

building_permits <- permits_issued %>%
    filter(building == 1) %>%
    group_by(m) %>%
    summarize(perc_in_target = sum(daystoissue <= 1)/n() * 100)

plot_data <- building_permits

ggplot(data = plot_data, aes(x = m, y = perc_in_target)) +
            theme(axis.text.x = element_text(angle = 45, hjust = 1),
                  plot.title = element_text(hjust = 0)) +
            geom_line(size = 1.5) +
            scale_x_yearmon(breaks = plot_data$m) +
            labs(color = NULL) +
            xlab("Month") +
            ylab("Percent") +
            ggtitle("Safety and Permits: Percentage of building permits issued within one day")

```

Source: LAMA; OPA database query

---

```{r sp_business_license_inspections}

license_inspections <- read_excel("business_license_inspections.xlsx")

names(license_inspections) <- tolower(names(license_inspections))

target <- 7

license_inspections <- license_inspections %>%
    mutate(m = as.yearmon(dateinspected),
           in_target = factor(ifelse(daystoinspect <= target, "Seven days or fewer", "More than seven days")))

license_inspections$in_target <- relevel(license_inspections$in_target, "Seven days or fewer")

plot_data <- license_inspections %>%
    group_by(m) %>%
    summarize(mean = mean(daystoinspect))

ggplot(data = plot_data, aes(x = m, y = mean)) +
    geom_line(size = 1.5) +
    geom_line(aes(y = target), linetype = 2) +
    scale_x_yearmon(breaks = plot_data$m) +
    labs(color = NULL) +
    theme(axis.text.x = element_text(angle = 45, hjust = 1), plot.title = element_text(hjust = 0)) +
    xlab("Month") +
    ylab("Days") +
    ggtitle("Safety and Permits: Mean days to complete business license zoning inspections")

```

Source: LAMA; OPA database query

---

```{r}

plot_data <- license_inspections

ggplot(data = plot_data, aes(x = m, fill = in_target)) +
    theme(axis.text.x = element_text(angle = 45, hjust = 1), plot.title = element_text(hjust = 0)) +
    geom_bar(position = "dodge", stat = "bin") +
    scale_x_yearmon(breaks = unique(plot_data$m)) +
    labs(fill = NULL) +
    xlab("Inspections") +
    ylab("Month") +
    ggtitle("Safety and Permits: Business license inspections by time elapsed from application")

```

Source: LAMA; OPA database query

---

```{r hdlc_permits_issued}

permits_issued <- read_excel("hdlc_permits_issued.xlsx")

names(permits_issued) <- tolower(names(permits_issued))

permits_issued <- permits_issued %>%
    mutate(m = as.yearmon(dateissued))

permits_issued <- permits_issued %>%
    group_by(m) %>%
    summarize(mean = mean(daystoissue), median = median(daystoissue))

ggplot(data = permits_issued, aes(x = m)) +
    geom_line(aes(y = mean, color = "Mean"), size = 1.5) +
    geom_line(aes(y = median, color = "Median"), size = 1.5) +
    scale_x_yearmon(breaks = permits_issued$m) +
    labs(color = NULL) +
    theme(axis.text.x = element_text(angle = 45, hjust = 1), plot.title = element_text(hjust = 0)) +
    xlab("Month") +
    ylab("Days") +
    ggtitle("Historic District Landmarks Commission: Days to issue staff-approvable permits")

```

Source: LAMA; OPA database query

---

```{r vcc_permits_issued}

permits_issued <- read_excel("vcc_permits_issued.xlsx")

names(permits_issued) <- tolower(names(permits_issued))

permits_issued <- permits_issued %>%
    mutate(m = as.yearmon(dateissued))

permits_issued <- permits_issued %>%
    group_by(m) %>%
    summarize(mean = mean(daystoissue))

ggplot(data = permits_issued, aes(x = m, y = mean)) +
    geom_line(size = 1.5) +
    scale_x_yearmon(breaks = permits_issued$m) +
    labs(color = NULL) +
    theme(axis.text.x = element_text(angle = 45, hjust = 1), plot.title = element_text(hjust = 0)) +
    xlab("Month") +
    ylab("Days") +
    ggtitle("Vieux Carre Commission: Mean days to issue staff-approvable permits")

```

Source: LAMA; OPA database query

## One Stop Shop

---

```{r}

oss <- read_excel("oss.xlsx")

names(oss) <- tolower(names(oss))
oss$queue <- tolower(oss$queue)

oss <- oss %>%
    filter(waittime >= 0 & waittime <= 480,
           servicetime >= 0 & servicetime <= 480) %>%
    mutate(m = as.yearmon(arrived))

# Consolidate queues into larger categories if needed
taxi_queues <- c("brake tag",
                 "bus permit",
                 "cpnc",
                 "driver/operator",
                 "taxi deputy director")

# Calculate performance measures

any_lic_perm_wait <- oss %>%
    filter(grepl("license|permit", queue)) %>%
    group_by(m) %>%
    summarize(med_wait_time = median(waittime), customers = n())

build_perm_wait <- oss %>%
    filter(queue == "building permit") %>%
    group_by(m) %>%
    summarize(med_wait_time = median(waittime), customers = n())

occ_lic_wait <- oss %>%
    filter(queue == "occup. license") %>%
    group_by(m) %>%
    summarize(med_wait_time = median(waittime), customers = n())

payment_wait <- oss %>%
    filter(queue == "payment") %>%
    group_by(m) %>%
    summarize(med_wait_time = median(waittime), customers = n())

taxi_wait <- oss %>%
    filter(queue %in% taxi_queues) %>%
    group_by(m) %>%
    summarize(mean_wait_time = mean(waittime), customers = n())

```

```{r}

target <- 12

plot_data <- any_lic_perm_wait

p1 <- ggplot(data = plot_data, aes(x = m, y = med_wait_time)) +
    geom_line(size = 1.5) +
    geom_line(aes(y = target), linetype = 2) +
    scale_x_yearmon(breaks = plot_data$m) +
    labs(color = NULL) +
    theme(axis.text.x = element_text(angle = 45, hjust = 1), plot.title = element_text(hjust = 0)) +
    xlab("Month") +
    ylab("Minutes") +
    ggtitle("Median wait time")

p2 <- ggplot(data = plot_data, aes(x = m, y = customers)) +
    theme(axis.text.x = element_text(angle = 45, hjust = 1), plot.title = element_text(hjust = 0)) +
    geom_bar(stat = "identity") +
    scale_x_yearmon(breaks = unique(plot_data$m)) +
    xlab("Month") +
    ylab("Customers") +
    ggtitle("Volume")

grid.arrange(p1, p2, ncol = 1, top = "One Stop Shop: Any license or permit queue")

```

Source: LobbyCentral; OPA database query

---

```{r}

target <- 18

plot_data <- build_perm_wait

p1 <- ggplot(data = plot_data, aes(x = m, y = med_wait_time)) +
    geom_line(size = 1.5) +
    geom_line(aes(y = target), linetype = 2) +
    scale_x_yearmon(breaks = plot_data$m) +
    labs(color = NULL) +
    theme(axis.text.x = element_text(angle = 45, hjust = 1), plot.title = element_text(hjust = 0)) +
    xlab("Month") +
    ylab("Minutes") +
    ggtitle("Median wait time")

p2 <- ggplot(data = plot_data, aes(x = m, y = customers)) +
    theme(axis.text.x = element_text(angle = 45, hjust = 1), plot.title = element_text(hjust = 0)) +
    geom_bar(stat = "identity") +
    scale_x_yearmon(breaks = unique(plot_data$m)) +
    xlab("Month") +
    ylab("Customers") +
    ggtitle("Volume")

grid.arrange(p1, p2, ncol = 1, top = "One Stop Shop: Building permit queue")

```

Source: LobbyCentral; OPA database query

---

```{r}

target <- 18

plot_data <- occ_lic_wait

p1 <- ggplot(data = plot_data, aes(x = m, y = med_wait_time)) +
    geom_line(size = 1.5) +
    geom_line(aes(y = target), linetype = 2) +
    scale_x_yearmon(breaks = plot_data$m) +
    labs(color = NULL) +
    theme(axis.text.x = element_text(angle = 45, hjust = 1), plot.title = element_text(hjust = 0)) +
    xlab("Month") +
    ylab("Minutes") +
    ggtitle("Median wait time")

p2 <- ggplot(data = plot_data, aes(x = m, y = customers)) +
    theme(axis.text.x = element_text(angle = 45, hjust = 1), plot.title = element_text(hjust = 0)) +
    geom_bar(stat = "identity") +
    scale_x_yearmon(breaks = unique(plot_data$m)) +
    xlab("Month") +
    ylab("Customers") +
    ggtitle("Volume")

grid.arrange(p1, p2, ncol = 1, top = "One Stop Shop: Occupational license queue")

```

Source: LobbyCentral; OPA database query

---

```{r}

target <- 5

plot_data <- payment_wait

p1 <- ggplot(data = plot_data, aes(x = m, y = med_wait_time)) +
    geom_line(size = 1.5) +
    geom_line(aes(y = target), linetype = 2) +
    scale_x_yearmon(breaks = plot_data$m) +
    labs(color = NULL) +
    theme(axis.text.x = element_text(angle = 45, hjust = 1), plot.title = element_text(hjust = 0)) +
    xlab("Month") +
    ylab("Minutes") +
    ggtitle("Median wait time")

p2 <- ggplot(data = plot_data, aes(x = m, y = customers)) +
    theme(axis.text.x = element_text(angle = 45, hjust = 1), plot.title = element_text(hjust = 0)) +
    geom_bar(stat = "identity") +
    scale_x_yearmon(breaks = unique(plot_data$m)) +
    xlab("Month") +
    ylab("Customers") +
    ggtitle("Volume")

grid.arrange(p1, p2, ncol = 1, top = "One Stop Shop: Payment queue")

```

Source: LobbyCentral; OPA database query

## Revenue

```{r}

revenue <- read_excel("revenue.xlsx")

names(revenue) <- tolower(names(revenue))
revenue$type <- tolower(revenue$type)

revenue <- revenue %>%
    filter(waittime >= 0 & waittime <= 480,
           servicetime >= 0 & servicetime <= 480) %>%
    mutate(m = as.yearmon(arrived))

# Calculate performance measures

abo <- revenue %>%
    filter(type == "abo") %>%
    group_by(m) %>%
    summarize(med_wait_time = median(waittime), customers = n())

accounts <- revenue %>%
    filter(type == "accounts") %>%
    group_by(m) %>%
    summarize(med_wait_time = median(waittime), customers = n())

administration <- revenue %>%
    filter(type == "administration") %>%
    group_by(m) %>%
    summarize(med_wait_time = median(waittime), customers = n())

business <- revenue %>%
    filter(type == "business") %>%
    group_by(m) %>%
    summarize(med_wait_time = median(waittime), customers = n())

enforcement_events <- revenue %>%
    filter(type == "enforcement/special events") %>%
    group_by(m) %>%
    summarize(med_wait_time = median(waittime), customers = n())

```

---

```{r}

plot_data <- abo

p1 <- ggplot(data = plot_data, aes(x = m, y = med_wait_time)) +
    geom_line(size = 1.5) +
    scale_x_yearmon(breaks = plot_data$m) +
    labs(color = NULL) +
    theme(axis.text.x = element_text(angle = 45, hjust = 1), plot.title = element_text(hjust = 0)) +
    xlab("Month") +
    ylab("Minutes") +
    ggtitle("Median wait time")

p2 <- ggplot(data = plot_data, aes(x = m, y = customers)) +
    theme(axis.text.x = element_text(angle = 45, hjust = 1), plot.title = element_text(hjust = 0)) +
    geom_bar(stat = "identity") +
    scale_x_yearmon(breaks = unique(plot_data$m)) +
    xlab("Month") +
    ylab("Customers") +
    ggtitle("Volume")

grid.arrange(p1, p2, ncol = 1, top = "Revenue: ABO queues")

```

Source: LobbyCentral; OPA database query

---

```{r}

plot_data <- accounts

p1 <- ggplot(data = plot_data, aes(x = m, y = med_wait_time)) +
    geom_line(size = 1.5) +
    scale_x_yearmon(breaks = plot_data$m) +
    labs(color = NULL) +
    theme(axis.text.x = element_text(angle = 45, hjust = 1), plot.title = element_text(hjust = 0)) +
    xlab("Month") +
    ylab("Minutes") +
    ggtitle("Median wait time")

p2 <- ggplot(data = plot_data, aes(x = m, y = customers)) +
    theme(axis.text.x = element_text(angle = 45, hjust = 1), plot.title = element_text(hjust = 0)) +
    geom_bar(stat = "identity") +
    scale_x_yearmon(breaks = unique(plot_data$m)) +
    xlab("Month") +
    ylab("Customers") +
    ggtitle("Volume")

grid.arrange(p1, p2, ncol = 1, top = "Revenue: Account queues")

```

Source: LobbyCentral; OPA database query

---

```{r}

plot_data <- business

p1 <- ggplot(data = plot_data, aes(x = m, y = med_wait_time)) +
    geom_line(size = 1.5) +
    scale_x_yearmon(breaks = plot_data$m) +
    labs(color = NULL) +
    theme(axis.text.x = element_text(angle = 45, hjust = 1), plot.title = element_text(hjust = 0)) +
    xlab("Month") +
    ylab("Minutes") +
    ggtitle("Median wait time")

p2 <- ggplot(data = plot_data, aes(x = m, y = customers)) +
    theme(axis.text.x = element_text(angle = 45, hjust = 1), plot.title = element_text(hjust = 0)) +
    geom_bar(stat = "identity") +
    scale_x_yearmon(breaks = unique(plot_data$m)) +
    xlab("Month") +
    ylab("Customers") +
    ggtitle("Volume")

grid.arrange(p1, p2, ncol = 1, top = "Revenue: Business queue")

```

Source: LobbyCentral; OPA database query

---

```{r}

plot_data <- enforcement_events

p1 <- ggplot(data = plot_data, aes(x = m, y = med_wait_time)) +
    geom_line(size = 1.5) +
    scale_x_yearmon(breaks = plot_data$m) +
    labs(color = NULL) +
    theme(axis.text.x = element_text(angle = 45, hjust = 1), plot.title = element_text(hjust = 0)) +
    xlab("Month") +
    ylab("Minutes") +
    ggtitle("Median wait time")

p2 <- ggplot(data = plot_data, aes(x = m, y = customers)) +
    theme(axis.text.x = element_text(angle = 45, hjust = 1), plot.title = element_text(hjust = 0)) +
    geom_bar(stat = "identity") +
    scale_x_yearmon(breaks = unique(plot_data$m)) +
    xlab("Month") +
    ylab("Customers") +
    ggtitle("Volume")

grid.arrange(p1, p2, ncol = 1, top = "Revenue: Enforcement and special events queue")

```

Source: LobbyCentral; OPA database query